England-India test Series, 2014
========================================================

This is an analysis of the peformance of the first 5 batsmen (not necessarily the "top" five) of each side.

```{r ,echo=FALSE}
#Read the data first. Note, we are using echo=FALSE here as we do not want to see
#the R code in the final output
scores<-read.csv("knitr_ex_1.csv")
#Split the data frame into subsets for India and England
#We could have done this using the subscript operator [],
#but will try the subset() function here
india<-subset(scores,side=="India")
england<-subset(scores,side=="England")
```
As a first step we add the total runs scored by these batsmen

```{r ,echo=FALSE}
#We could have used our familiar ddply() as follows: 
#ddply(india,"batsman", function(y) sum(y$runs))
#Trying out aggretage() which is part of the base R package
india.total.runs<-aggregate(india["runs"],by=list(batsman=india$batsman),FUN=sum) #india[,5] is also OK instead of "india["runs"]" but the latter is probably more intuituve
england.total.runs<-aggregate(england["runs"],by=list(batsman=england$batsman),FUN=sum)
#Sort them in descending order

india.total.runs<-india.total.runs[order(india.total.runs$runs, decreasing=TRUE),]
england.total.runs<-england.total.runs[order(england.total.runs$runs, decreasing=TRUE),]
#Having obtained the summaries, we print them in tabular form in the next chunk
```
As we can see `r as.character(india.total.runs[1,1])` has scored themost runs for India so far
```{r , results='asis', message=FALSE, echo=FALSE} 
require(knitr)
kable(india.total.runs, format="html")
```
while `r as.character(england.total.runs[1,1])` piled the maximum runs for England
```{r , results='asis', message=FALSE, echo=FALSE} 
kable(england.total.runs, format="html")
```

We would now look at the averages per match and the strike rates, which would complete the picture of how good each of these batsmen were

```{r ,echo=FALSE}
india.avg.runs<-aggregate(india["runs"],by=list(batsman=india$batsman),FUN=mean)
england.avg.runs<-aggregate(england["runs"],by=list(batsman=england$batsman),FUN=mean)
india.avg.runs<-india.avg.runs[order(india.avg.runs$runs, decreasing=TRUE),]
england.avg.runs<-england.avg.runs[order(england.avg.runs$runs, decreasing=TRUE),]

```
For India, `r as.character(india.avg.runs[1,1])` had the best per-match average
```{r , results='asis', message=FALSE, echo=FALSE} 
require(knitr)
kable(india.avg.runs, format="html")
```
whereas `r as.character(england.avg.runs[1,1])` shone for England with the best average:

```{r , results='asis', message=FALSE, echo=FALSE} 
require(knitr)
kable(england.avg.runs, format="html")
```
What about strike rates ? Lets look at the runs scored per 100 deliveries. 

```{r,message=FALSE,echo=FALSE}
#Let's use our new-found knowledge of transform() in the plyr package to 
#add a "calculated"" field "strike_rate"
require(plyr)
india<-ddply(india,"batsman", transform, strike_rate=(runs/balls_Faced)*100.0)
england<-ddply(england,"batsman", transform, strike_rate=(runs/balls_Faced)*100.0)

india.strike.rate<-aggregate(india["strike_rate"],by=list(batsman=india$batsman),FUN=mean)
england.strike.rate<-aggregate(england["strike_rate"],by=list(batsman=england$batsman),FUN=mean)
india.strike.rate<-india.strike.rate[order(india.strike.rate$strike_rate, decreasing=TRUE),]
england.strike.rate<-england.strike.rate[order(england.strike.rate$strike_rate, decreasing=TRUE),]

```
For India, `r as.character(india.strike.rate[1,1])` had the best strike rate
```{r , results='asis', message=FALSE, echo=FALSE} 
kable(india.strike.rate, format="html")
```
whereas `r as.character(england.strike.rate[1,1])` had the best strike rate for England:

```{r , results='asis',echo=FALSE} 
kable(england.strike.rate, format="html")
```

Let's see how the strike rates stackup graphically:

```{r fig.width=7, fig.height=6, echo=FALSE, message=FALSE}
plot(india.strike.rate$strike_rate, col="red", type="l", pch=19, ylab="Strike Rates", xlab="Batsmen")
points(england.strike.rate$strike_rate, col="green", type="l", pch=19)
legend("topright", c("India", "England"), lty=c(1,1), col=c("red","green"))
```
Let's see how these batsmen have got out. As one would expect in England - where the ball swings a lot - one would expect a lot of them to be out caught. We will find out if that's indeed the case

```{r, echo=FALSE, message=FALSE}
#Use the table() function to create a frequency table and sort them in descending order
india.outs<-table(india$how_out)
india.outs<-india.outs[order(india.outs, decreasing=T) ]
england.outs<-table(england$how_out)
england.outs<-england.outs[order(england.outs, decreasing=T) ]
```
As expected `r as.character(labels(india.outs[1][1]))` comprises the biggest chunk of dismissals of the Indian batsmen whereas `r as.character(labels(england.outs[1][1]))` was the reason for sending the English batsmen on the long walk back to thw pavilion. We can create a pie to compare how the categories stack up

```{r , message=FALSE, echo=FALSE,fig.width=10, fig.height=9}
#First create a plot area to hold two pie charts in a row
par(mfrow=c(1,2))
pie(india.outs, main="India - Dismissals")
pie(england.outs, main="England - Dismissals")
```
